A Test of the Frequency Independence Assumption of the Power System Parameters used in Geomagnetically Induced Current Calculations

### 2019-02-14 paper.tex
Updates from Overleaf

\begin{keypoints}
\item The power system parameters derived from GIC and geoelectric field measurements are frequency dependent.
\item A GIC prediction model with frequency dependence significantly outperforms a frequency independent model.
\item The best predictions are found when the model input is not the measured geoelectric field but rather the geoelectric field predicted by a transfer function model driven by the geomagnetic field.
\end{keypoints}

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory \todo[color=blue!40]{Shinichi: Is 10 km correct?} $\sim${\color{red}\textbf{10}} kilometers away. We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

@@

Geomagnetically induced currents (GICs) are electric currents in man-made conducting systems due to electric fields induced near Earth's surface. These fields are a result of time variations in ionospheric currents on timescales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}. Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}. Accurate estimation of GICs given direct measurements of the electric field (or estimates based on the geomagnetic field) is important for power system design, retrospective analysis, and for the mitigation of the impacts of space weather on power systems \todo{Pierre and Trevor - any suggestions?} {\color{red} \textbf{[REFS]}}.

A common assumption made in estimates of GIC using either measured or estimated values of the horizontal geoelectric field at Earth's surface, $\mathbf{E}$, is that the power system is resistive or quasi-DC so that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}.  The coefficients $a_o$ and $b_o$ can then be calculated analytically using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}. This assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.

@@

The 1-second-cadence GIC data used in this paper span a total of 35 days across the 10 intervals: 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22 that have been presented previously in \citet{Watari2009} and \cite{Watari2015}. This represents all of the data that were available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements.

One-second-cadence geoelectric field and geomagnetic field data from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}. Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.

Spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$. In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$. For all of the time intervals considered, the fraction of interpolated $\mathbf{E}$ values ranged from 0--0.1\%.

The only modifications made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values. The values for $E_x$ and $E_y$ were the same and much larger than the surrounding values.

The GIC dataset includes 1-second raw data and 1 Hz low-pass filtered raw measurements.  The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{sample}. For all of the time intervals considered, the fraction of interpolated GIC values ranged from 1--4\%.

The GIC data contains non-physical spikes that are followed by a relaxation time of  40-80 seconds. These spikes were removed by identifying times $t_s$ when the change in 1-second $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$. In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.

\section{Models}

To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model~1,

@@

\noindent
the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $GIC$ is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations. The values of $\mathbf{p}$ are then obtained using a standard linear regression solver. The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ for all values of $t$. \citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}

In Model 2, the system parameters are frequency dependent

\begin{linenomath*}
\begin{equation}
G_E(\omega) = A(\omega)E_x(\omega) + B(\omega)E_y(\omega)
\label{model2}
\tag{2a}
\end{equation}
\end{linenomath*}

Note that the scalar quantity $B$ without a subscript refers to the system parameter in Model~2 while the vector quantity $\mathbf{B}$ or its components, $B_x$ and $B_y$ refer to the magnetic field.

The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:

\setcounter{equation}{2}
\begin{linenomath*}
\begin{equation}
G_{E'}(\omega) = A(\omega)E'_x(\omega) + B(\omega)E'_y(\omega)
\end{equation}
\end{linenomath*}

@@

\begin{equation*}
G_{E'}(\omega) = \big[A(\omega)Z_{xx}(\omega) + B(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ A(\omega)Z_{xy}(\omega) + B(\omega)Z_{yy}(\omega) \big] B_y(\omega)
\end{equation*}

That is, instead of using the measured electric field directly as in Model 2, we use a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ to provide an estimate of the electric field based on the magnetic field. In this work, we have computed $\mathbf{Z}$ using the method described later in this section.

A fourth model considered is:

\begin{linenomath*}
\begin{equation}
G_B(\omega) = Z_x(\omega)B_x(\omega) + Z_y(\omega)B_y(\omega)
\end{equation}
\end{linenomath*}

\noindent
In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and the model is expected to produce predictions of $GIC$ that are similar to that obtained from Model 3.

For frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$. To convert the frequency domain transfer functions $A(\omega)$ and $B(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed. The transfer function magnitudes $|A(\omega)|$ and $|B(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude). The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$. Additional details on the method are given in \cite{Weigel2017} and references therein.

@@

Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data. The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components of the 35 transfer functions. The average impulse response functions were obtained by averaging the 35 impulse response functions.

There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic least-squares method \citep{Simpson2005}. These works typically involve the estimation of a robust and bias-corrected transfer function. The results presented used were obtained using the basic least-squares transfer function estimation method as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters. As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).

However, several alternative methods for computing the transfer functions were tested individually to determine if there was a change in the computed transfer functions and, most important for this work, each model's out-of-sample prediction ability. When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance are within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$ seconds. The experiments in this regard are discussed in the following section and in the Appendix.

@@

The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as $PE$ = 1 - variance(Prediction Error)/variance(Predicted) and is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.

The average PE and MSE ratio was computed by averaging the 35 PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the 35 signal-to-noise ratios at that frequency. The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the 35 prediction metrics. The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.

@@

Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 4 has the highest PE and Model 1 has the lowest.

@@

Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models. The error bars extend by $\pm$ one standard error of the 35 SN ratios at each period. The SN ratio for Model 1 is always lower than that for Models 2-4.  Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1. Below $500$~s (8~min), Model 2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} - at approximately 2006-08-07 07:00, the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.

The average parameters for Model 1 are $a_o = $ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~2. (Results are similar if 1-minute averages are used; \cite{Watari2015} found \todo[color=blue!40]{Shinichi: I get $a_o=45.8$ and $b_o=-59.9$ in this time interval.} $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).

The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of GIC \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GIC, but is a poor approximation relative to Models 2 and 3. For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.

@@

Figure \ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar and are different from zero at more than one time value (the impulse response for Model 3 is similar to that of Model 2 and is not shown). By definition, the impulse response for Model 1 is zero except at $t=0$. Note that the sign of the impulse response at $t=0$ is opposite for Model 1 and Model 2 for the $E_y$ coefficient - at $t=0$, $b_o < 0$ while $b(0)>0$.

Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition). Note that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2. In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}) so that $E_x$ and $E_y$ on average contribute approximately equal amounts to the predictability of GIC.

Figure~\ref{Phi} shows the transfer function phases. By definition, that for Model 1 is zero. The phases for $Z_x$ and $B$ are offset by 120 degrees \todo{Why?} while the phases for $A$ and $B$ are significantly different \todo{Why?}.

Although the primary objective of this work is the creation of the models for comparison of the predictive ability, a natural outcome is model parameter estimates. As noted in the introduction, the motivation for using a basic least squares method is (1) that it reduces the number of adjustable methodological parameters that must be considered and (2) because we are attempting to consider which model performs best in a least-square-error sense. It is natural to ask if the model parameter estimates are biased and how sensitive they depend on the statistical method used. Several alternative estimation methods have been considered and are discussed in the Appendix; the primary conclusion is that the values shown in Table~1 are not sensitive and the parameter estimates are only sensitive below approximately $T=30$~s (they are unreliable in the sense that the alternative estimation methods give parameter values outside of the error bars shown for the baseline method).

@@

Because Model~4 provided the best prediction performance, the ideal measurements to use for estimating GIC in this dataset are contemporaneous GIC and $\mathbf{B}$ measurements.

@@

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}. GIC data provided by ... \todo[color=blue!40]{Shinichi: How should this be referenced?}

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
\caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
\label{SN}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
\caption{Distribution of parameters in Model 1. Each set of parameters were computed 35 times using one day of 1-second cadence data. The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
\label{histogram}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_H-options-1.pdf}
\caption{Impulse responses for Models 1, 2, and 4. The circles at $t = 0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4. The values of $a_o$ and $b_o$ have been scaled by a factor of 10 for readability.}
\label{H}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_Z-options-1.pdf}
\caption{Frequency domain transfer functions for Models 1, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
\label{Z}
\end{figure}

>>>>

\begin{keypoints}
\item The power system parameters derived from GIC and geoelectric field measurements are frequency dependent.
\item A GIC prediction model with frequency dependence significantly outperforms a frequency independent model.
\item A model derived using GIC/geomagnetic field data provides better predictions than a GIC/geoelectric field-based model.
\end{keypoints}

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10\todo[color=blue!40]{Shinichi: Is 10 km correct?} kilometers away.  We compute frequency dependent network parameters $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

@@

Geomagnetically induced currents (GICs) are electric currents in conducting systems due to electric fields induced near Earth's surface. These fields are a result of time variations in ionospheric currents on timescales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}. Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}. Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \todo{Pierre and Trevor - any suggestions?} {\color{red} \textbf{[REFS]}}.

A common assumption made in estimates of GIC using either measured or estimated values of the horizontal geoelectric field at Earth's surface, $\mathbf{E}$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}. In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression, when contemporaneous measurements of GIC and $\mathbf{E}$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.][]{Boteler2014a,Boteler2014b}. This quasi-DC assumption has been used or stated in many works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Viljanen2012,Overbye2012}.

@@

The 1-second-cadence GIC data used span a total of 35 days across 10 intervals: 2006-04-03 -- 2006-04-10, 2006-07-25 -- 2006-07-29, 2006-08-05 -- 2006-08-08, 2006-08-19 -- 2006-08-21, 2006-11-07 -- 2006-11-12, 2006-11-28 -- 2006-11-29, 2006-12-01, 2006-12-14 -- 2006-12-15, 2007-11-18 -- 2007-11-20, and 2007-11-22 -- 2007-11-22. These data have been presented previously in \citet{Watari2009} and \cite{Watari2015}. These time intervals cover those for which GIC data was made available with the exception of three days for which there were obvious problems with either the electric field or GIC measurements. The GIC dataset includes 1-s raw data and 1 Hz low-pass filtered raw measurements. The results presented here are for the 1 Hz low-pass filtered GIC values

One-second-cadence $\mathbf{E}$ and $\mathbf{B}$ measurements a from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal. Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.

Data for the longest continuous time span of GIC measurements, 2006-04-02 through 2006-04-10 are shown Figure~\ref{sample}.

Spikes in the electric field measurements were assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$. In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$. For all of the time intervals considered, the fraction of interpolated $\mathbf{E}$ values ranged from 0--0.1\%.

The only modification made to any of the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values. The values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.

The GIC data contains non-physical spikes that are followed by a relaxation time of 40-80 s. These spikes were removed by identifying times $t_s$ when the change in 1-se $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$. In the example interval shown in Figure~\ref{sample}, 493 spikes were removed. For all of the time intervals considered, the fraction of interpolated GIC values ranged from 1--4\%.

\section{Models and Methods}

To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model ~1,

@@

\noindent
the matrix equation $\mathbf{GIC} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\mathbf{GIC}$ is a 864000x1 matrix of 1 day of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations. The values of $\mathbf{p}$ are then obtained using a standard linear regression solver. The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval. \citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}. The parameters $a_o$ and $b_o$ were computed for each of the 35 unique 1-day window segments of available data.

In Model 2 the system parameters are frequency dependent

\begin{linenomath*}
\begin{equation}
G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)\,.
\label{model2}
\end{equation}
\end{linenomath*}

The third model considered uses an electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$:

\setcounter{equation}{2}
\begin{linenomath*}
\begin{equation}
G_{E'}(\omega) = a(\omega)E'_x(\omega) + b(\omega)E'_y(\omega)
\label{model3}
\end{equation}
\end{linenomath*}

@@

\begin{equation*}
G_{E'}(\omega) = \big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)\,.
\end{equation*}

That is, instead of using the measured electric field directly as in Model 2, a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field. In this work, $\boldsymbol{Z}$ was calculated using the method described later in this section; $a(\omega)$ and $b(\omega)$ are taken from the solution of Equation~\ref{model2}, using the method also described later in this section.

The fourth model is

\begin{linenomath*}
\begin{equation}
G_B(\omega) = z_x(\omega)B_x(\omega) + z_y(\omega)B_y(\omega)\,.
\label{model4}
\end{equation}
\end{linenomath*}

\noindent
In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of GIC and $\mathbf{B}$ and the model is expected to produced predictions of GICs that are similar to that obtained from Model 3.

Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e = 2\pi f_e$, by performing a least squares regression using Equation~\ref{model2} and the set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the fourier transform of 24 hours of 1-second data. The values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$; if $f_e$ value is not a multiple of $1/86400$, the nearest multiple is chosen. The regression for $f_e$ uses all $f$ values between the next higher and lower value of $f_e$.

To convert the frequency domain transfer functions $a(\omega)$ and $b(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed. The transfer function magnitudes $|a(\omega)|$ and $|b(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude). The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}. For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$, the $xx$ and $yx$ components of $\boldsymbol{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$; the $yx$ and $yy$ components are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$.

@@

Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data. The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components, respectively, of the 35 transfer functions. The average impulse response functions were obtained by averaging the 35 impulse response functions.

There are many statistical methods that have been developed for estimating transfer functions in the magnetotelluric literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}. These works typically involve the estimation of a robust and bias-corrected transfer function. In the results presented, we did not use a robust statistical method or a bias correction algorithm as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters. As discussed in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).

However, several alternative methods were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and, most important for this work, the out-of-sample prediction performance of each model. When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance is within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$~s). The experiments in this regard are discussed in the Appendix.

@@

The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted, and it represents the fraction of the variance in the predicted data that predicted by the model.

The average PE and MSE ratio was computed by averaging the 35 out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample 35 signal-to-noise ratios at that frequency. The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the 35 prediction metrics. The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.

@@

Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.

@@

Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models. The error bars represent $\pm$ one standard error of the 35 SN ratios at each period. Consistent with the results in Table~1, The SN ratio for Model 1 is lower than that for Models 2-4 at all frequencies.  Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio. Below $500$~s (8~min), Model 2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.

The average parameters for Model 1 are $a_o = $ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found \todo[color=blue!40]{Shinichi: I get $a_o=45.8$ and $b_o=-59.9$ in this time interval.} $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).

The wide distribution and large error bars are suggestive that Model~\ref{model1} is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$  \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3. For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.

@@

Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar and are different from zero at more than one time value (the impulse response for Model 3 is similar to that of Model 2 and is not shown). By definition, the impulse response for Model 1 is zero except at $t=0$. Note that the sign of the impulse response at $t=0$ is opposite for Model 1 and Model 2 for the $E_y$ coefficient - at $t=0$, $b_o < 0$ while $b(0)>0$.

Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition). Note that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2. In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).

Figure~\ref{Phi} shows the transfer function phases. By definition, that for Model 1 is zero. The phases for $z_x$ and $b$ are offset by 120 degrees \todo{Why?} while the phases for $a$ and $b$ are significantly different \todo{Why?}.

@@

@@

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}. \todo[color=blue!40]{Shinichi: How should this be referenced?} GIC data provided by ...

%\clearpage

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summary_SN-options-1.pdf}
\caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
\label{SN}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summary_aobo_histograms-options-1.pdf}
\caption{Distribution of parameters in Model 1. Each set of parameters were computed 35 times using one day of 1-second cadence data. The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
\label{histogram}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summary_H-options-1.pdf}
\caption{Impulse responses for Models 1, 2, and 4. The circles at $t = 0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4. The values of $a_o$ and $b_o$ have been scaled by a factor of 10 for readability.}
\label{H}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summary_Z-options-1.pdf}
\caption{Frequency domain transfer functions for Models 1-3, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
\label{Z}
\end{figure}
